Comparing probabilistic predictive models applied to football

### 2015-08-22 template.tex
com comentarios do Danilo

	Figure \ref{fig::entropy} presents the boxplots of the entropy of each prediction for the predictive models considered. Recall
	that the entropy of a prediction $(p_1,p_2,p_3)$ is given by $- \sum_{i=1}^3 p_i \log{p_i}$.
	Since the entropy values are smaller than 1.09, the plots show that the predictions
	are typically more informative than the trivial prediction. Nevertheless, all methods yielded similar entropies in average, that is,
	none of them seem to provide more informative probabilities.

@@

	%{\color{darkgreen}Therefore, our analyses show that, marginally, all methods yield similar scores in average, i.e., they have similar predictive power. A related question
	%is whether, for each match, the performance of the methods is also similar. Figure \ref{fig::scores2} displays pairwise comparisons
	%of the various score between the five models we investigate. All methods present
	%a reasonable agreement among each other. Evidently, the Multinomial-Dirichlet models
	%are the ones where there is largest agreement. To formally test the hypotheses that the models yield the same scores in average for
	%each match,
	%we use Friedman's rank sum test. All p-values are not significant at the level 5\% (Brier: 0.077; Spherical: 0.055; Proportion
	%of errors: 0.655), except for the logarithmic
	%score, which has p-value of 0.021. Post-hoc analyses show that the differences are in Mn-Dir$_1$ versus BB2 (p=0.031),
	%Mn-Dir$_2$ versus BB1 (p=0.043), and Mn-Dir$_3$ versus BB1 (p=0.030). These results thus indicate that BB1 and BB2
	%are similar to each other in terms of predictive power, while the performance of Mn-Dir models differ from the BB's \emph{for each match}.
	%We therefore conclude that while the models we investigate may yield different scores for each
	%match according to some metrics, the overall scores are similar. That is, while in some matches Mn-Dir models
	%have better results, for others BB models have better performance. However, in the long term, all methods have similar accuracy.
	%}
	{\color{darkgreen}Therefore, our analyses show that all methods yield similar scores in average, i.e., they have similar predictive power. A related question
	is whether, for each match, the performance of the methods is also similar. To answer this question we compared the models pairwise, and the result is found on Figure \ref{fig::scores2}, where the plots display the scores of each match for every couple of models considered. The plots show that all methods are reasonably similar and, evidently, the Multinomial-Dirichlet models
	are the ones that agree the most. To formally test the hypotheses that the models yield the same scores in average for each match, we used Friedman's rank sum test. Except for the logarithmic score (p-value 0.021), all the other metrics did not present significant differences at the level of confidence of 5\% (Brier: 0.077; Spherical: 0.055; Proportion of errors: 0.655). Post-hoc analyses show that the differences are in Mn-Dir$_1$ versus BB2 (p=0.031),
	Mn-Dir$_2$ versus BB1 (p=0.043), and Mn-Dir$_3$ versus BB1 (p=0.030). These results thus indicate that BB1 and BB2
	are similar to each other in terms of predictive power, while the performance of Mn-Dir models differ from the BB models \emph{for each match}.
	We therefore conclude that while the models we investigate may yield different scores for each
	match according to some metrics, the overall scores are similar, i. e., while in some matches Mn-Dir models
	have better results, for others BB models have better performance. However, for a large set of matches, all methods have similar accuracy.
	}

		\begin{figure}[H]
			\centering
			\includegraphics[page=13,scale=0.48]{futebolComparacaoModelosForPaper.pdf}
			\includegraphics[page=14,scale=0.48]{futebolComparacaoModelosForPaper.pdf}\\
			\caption{Pairwise comparison of the various scores. Left plot: upper right plots display Brier Scores;
			lower left plots display Logarithmic Scores. Right plot: upper right plots display proportion of agreements
			between methods; lower left  plots display Spherical Scores. Red line represents the identity $x=y$.}
			\label{fig::scores2}
		\end{figure}

@@

	{\color{darkgreen}Therefore, we once more highlight that 
	although the accuracy of each method may depend on the specific characteristics of each match,
	all methods have similar predictive power, and none of them is more informative  in average. 
    Moreover, all methods yielded better predictions than the trivial prediction, and they were all well-calibrated.}

@@

		If, on the one hand, our findings appear as another supportive example of Ockham's razor, on the other they pose several questions about probabilistic prediction of sport events.
		In particular, based on the fact that the models have similar {\color{darkgreen}predictive power in average}, one may ask: is there an irreducible ``randomness'' or ``degree of unpredictability'' implicit in these events?
		Is this degree an indicator of how tight or leveled is the championship being studied?

>>>>

		%{\color{darkgreen}Therefore, our analyses show that, marginally, all methods yield similar scores in average, i.e., they have similar predictive power. A related question
		%is whether, for each match, the performance of the methods is also similar. Figure \ref{fig::scores2} displays pairwise comparisons
		%of the various score between the five models we investigate. All methods present
		%a reasonable agreement among each other. Evidently, the Multinomial-Dirichlet models
		%are the ones where there is largest agreement. To formally test the hypotheses that the models yield the same scores in average for
		%each match,
		%we use Friedman's rank sum test. All p-values are not significant at the level 5\% (Brier: 0.077; Spherical: 0.055; Proportion
		%of errors: 0.655), except for the logarithmic
		%score, which has p-value of 0.021. Post-hoc analyses show that the differences are in Mn-Dir$_1$ versus BB2 (p=0.031),
		%Mn-Dir$_2$ versus BB1 (p=0.043), and Mn-Dir$_3$ versus BB1 (p=0.030). These results thus indicate that BB1 and BB2
		%are similar to each other in terms of predictive power, while the performance of Mn-Dir models differ from the BB's \emph{for each match}.
		%We therefore conclude that while the models we investigate may yield different scores for each
		%match according to some metrics, the overall scores are similar. That is, while in some matches Mn-Dir models
		%have better results, for others BB models have better performance. However, in the long term, all methods have similar accuracy.
		%}
		{\color{darkgreen}Therefore, our analyses indicate that all methods yield similar scores in average, i.e., they have similar predictive power. In order to formally check this, 
		we test the hypotheses that all five models have the same median score.
		We do this by using Friedman's rank sum test for each score, a statistical test that takes into account
		the dependency between the observations (notice each match is evaluated by each model). 
		Except for the logarithmic score (p-value 0.021), all the other metrics did not present significant differences at the level of confidence of 5\% (Brier: 0.077; Spherical: 0.055; Proportion of errors: 0.655). Post-hoc analyses\footnote{We used
		Friedman tests for the pairwise comparisons as well.} for the logarithmic score show that the differences are in Mn-Dir$_1$ versus BB2 (p=0.031),
				Mn-Dir$_2$ versus BB1 (p=0.043), and Mn-Dir$_3$ versus BB1 (p=0.030). 
				The median logarithmic scores are 
				 0.97 for Mn-Dir$_1$, 0.94 for Mn-Dir$_2$, 0.95 for Mn-Dir$_3$,  0.97 for BB1, and  1.00
				 for BB2, thus indicating that the new models are slightly better according to this criterium (but not according to
				 the others).
                 We further illustrate this point in Figure \ref{fig::scores2}, where the plots display the scores of each match for every couple of models considered. The plots show that all methods are reasonably similar and, evidently, the Multinomial-Dirichlet models are the ones that agree the most. 
		}

			\begin{figure}[H]
				\centering
				\includegraphics[page=13,scale=0.48]{futebolComparacaoModelosForPaper.pdf}
				\includegraphics[page=14,scale=0.48]{futebolComparacaoModelosForPaper.pdf}\\
				\caption{Pairwise comparison of the various scores. Left plot: upper right plots display Brier Scores;
				lower left plots display Logarithmic Scores. Right plot: upper right plots display proportion of agreements
				between methods (i.e., proportion of times the conclusions are the same; see Section \ref{sec::calib}); lower left  plots display Spherical Scores. Red line represents the identity $x=y$.}
				\label{fig::scores2}
			\end{figure}

	In order to check how informative the models are, Figure \ref{fig::entropy} presents the boxplots of the entropy of each prediction for the predictive models considered. Recall
	that the entropy of a prediction $(p_1,p_2,p_3)$ is given by $- \sum_{i=1}^3 p_i \log{p_i}$.
	Since the entropy values are smaller than 1.09, the plots show that the predictions
	are typically more informative than the trivial prediction. Nevertheless, all methods yielded similar entropies in average, that is,
	none of them seem to provide more informative probabilities.

@@

@@

	{\color{darkgreen}Therefore, we once more highlight that  
	all methods have similar predictive power, and none of them is more informative  in average. 
    Moreover, all methods yielded better predictions than the trivial prediction, and they were all well-calibrated.}

@@

		If, on the one hand, our findings appear as another supportive example of Ockham's razor, on the other they pose several questions about probabilistic prediction of sport events.
		In particular, based on the fact that the models have similar predictive power in average, one may ask: is there an irreducible ``randomness'' or ``degree of unpredictability'' implicit in these events?
		Is this degree an indicator of how tight or leveled is the championship being studied?
